---
title: "Lab 8 - Linear Regression in R"
author: "Jason Johns"
date: "11/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# Load tidyverse & data

library(tidyverse)
salmon <- read_csv("salmon_imports.csv")

```


####4. Exploratory scatterplot
```{r}
# Exploratory scatterplot

salmon_scatter <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point()

salmon_scatter

# It seems like linea regression is an appropriate method for analyzing the relationship between year and fresh salmon imports. I have no major concerns about using regression for this analysis.

```


####5. Linear regression
```{r}
# Running regression

salmon_model <- lm(salmon_fresh ~ year, data = salmon)
salmon_model

# imports(tons) = -14982940 + 7550(year)
```

####  - The slope means that for every successional year, there is a 7550 ton increase in salmon imports.
####  - The largely negative y-intercept means that if we were to go back to 1989, there would be -14982940 tons of salmon imports. Thus, this model cannot be extrapolated to past values.


####6. Model diagnostics
```{r}
par(mfrow = c(2,2))
plot(salmon_model)

```

#### -Based on the qq plot, the residuals appear to be relatively normally distributed, other than at lower and upper theoretical quantiles
#### -There are concerns about some major outliers as well as a lack of heteroscedasticity. The residuals appear to be more widely distributed at larger fitted values.


####7. Explore model fit and significance
```{r}
# Looking further at model with summary()

summary(salmon_model)


```

#### - Year is a statistically significant predictor of salmon imports (p < 0.001).
#### - Based on the R^2 value of 0.9478, 94.8% of the variance in tons of salmon imports is predictable by year.
#### - 5.2% of the variance in tons of salmon imports is not predictable by year.

####8. Prepare a final graph
```{r}
# Prepare final graph

salmon_final_graph <- ggplot(salmon, aes(x = year, y = salmon_fresh)) + 
  geom_point() +
  geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") + theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) +
  labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports\n(1,000 pounds)")

salmon_final_graph

```

